---
title: "Analysis of the efficiency of MC-MoCo"
author: L. Soustelle, J. Lamy, A. Le Troter, M. Guye, J.-P. Ranjeva, F. Munsch, G.
  Varma, D. Alsop, J. Pelletier, O. Girard, G. Duhamel
date: ''
---

```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)
library(knitr)
library(multcomp)
library(nlme)
```

This document contains the statistical analysis described in *A Motion Correction Strategy for Multi-Contrast based 3D parametric imaging: Application to Inhomogeneous Magnetization Transfer (ihMT)*, by L. Soustelle et al., to appear in *Magnetic Resonance in Medicine*. The source code of this document, and its related data, is available at https://github.com/lamyj/mcmoco-data/.

The data contains in each row contains the position in physical space of one marker of one ihMT-weighted image of one subject, placed either before motion-correction (*Pre-MoCo*) or after application of a motion-correction algorithm (*MC-MoCo* or *MCFLIRT*).

```{r}
positions = read.csv("positions.csv", colClasses=c(Volume="factor"))
kable(head(positions))
```

# Quantification of the residual motion

For each subject and each set of images (uncorrected, MCFLIRT- and MC-MoCo-corrected), a reference point of each landmark is defined as the centroid of the landmark locations across the images of the set.

```{r}
references = aggregate(
    cbind(x, y, z) ~ Subject + Marker + Method, positions, mean)
```

The reference points are used to quantify the amount of native motion (uncorrected images) and residual motion (MCFLIRT- and MC-MoCo-corrected images) by averaging the distances between the reference points and their corresponding landmarks across all the subjects.

```{r}
df = merge(
    positions, references, by=c("Subject", "Marker", "Method"), 
    suffixes=c("", "_c"))
df$Distance = sapply(
    1:nrow(df), 
    function(i) { with(df[i,], norm(c(x, y, z) - c(x_c, y_c, z_c), "2")) })
distances = aggregate(Distance ~ Subject + Marker + Method, df, mean)
```

A boxplot of the native motion evaluated on the images before motion correction and of residual motion evaluated after application of MCFLIRT and MC-MoCo show a lower residual motion for MC-MoCo compared to that of MCFLIRT for all landmarks. The residual motion value is additionally rather constant across all markers for MC-MoCo.

```{r echo=FALSE, message=FALSE}
# Re-order factor to improve readability of plot.
distances$Method = factor(distances$Method, c("Pre-MoCo", "MCFLIRT", "MC-MoCo"))
distances$Marker = factor(
    distances$Marker, c("R-SFG", "R-PCG", "L-MFG", "LR-SCCG"))

plot = (
    ggplot(distances, aes(Method, Distance, color=Marker)) +
    geom_boxplot()+ylab("Native/Residual motion (mm)"))
ggsave("motion.png", plot)
print(plot)
```

# Statistical analysis

To evaluate the performance of MC-MoCo and MCFLIRT, and to verify that the behavior of both methods is constant across landmarks and between the two groups of subjects, we use a mixed-effect linear model with fixed effects for the motion correction method, the landmark, and the subject group (deduced from the subject ID), and with a random effect for the subject to account for the non-independence of the observations within a subject.

```{r}
# Deduce Group factor from subject ID
distances$Group = as.factor(
    sapply(distances$Subject, sub, pattern="(.*)_.*", replacement="\\1"))
# Re-order to improve readability of hypothesis testing.
distances$Method = factor(distances$Method, c("MC-MoCo", "MCFLIRT", "Pre-MoCo"))
model = lme(Distance ~ Method + Marker + Group, random=~1|Subject, distances)
```

The residuals of the mixed-effects linear model show neither deviation from a normal distribution (Shapiro-Wilk test, p=`r format.pval(shapiro.test(residuals(model))$p.value, digits=3)`) nor apparent structure in the variance of the residuals, thus validating the model.

```{r echo=FALSE}
info = data.frame(Residuals=residuals(model), Fitted=fitted(model))
qq_plot = ggplot(info, aes(sample=Residuals)) + geom_qq() + geom_qq_line()
residuals_plot = ggplot(info, aes(Fitted, Residuals)) + geom_point()
grid.arrange(qq_plot, residuals_plot, ncol=2)
```

The analysis of the individual factors shows a significant effect of the motion-correction methods, indicating a difference between the non-corrected images and the motion-corrected images, but no difference either across the landmarks or between the two groups of subjects, thereby indicating that differences in residual motions depends only on the algorithm.

```{r echo=FALSE}
df = as.data.frame(anova(model))
df$`p-value` = format.pval(df$`p-value`, digits=3, eps=1e-10)
kable(df, digits=3, align="lrrrr")
```

Further linear hypotheses tests comparing the native motion from residual motions derived from MCFLIRT and MC-MoCo show statistically significant reductions for MCFLIRT and for MC-MoCo as well as signficant reduction of MC-MoCo compared to MCFLIRT.

```{r}
hypotheses = glht(model, linfct=mcp(Method="Tukey"))
```

```{r echo=FALSE}
comparisons = summary(hypotheses)
# From the source of multcomp
pq = comparisons$test
mtests = data.frame(
    pq$coefficients, pq$sigma, pq$tstat, format.pval(pq$pvalues, eps=1e-10),
    stringsAsFactors=FALSE)
letter = ifelse(comparisons$df == 0, "z", "t")
pname = switch(
    comparisons$alternativ,
    less=paste0("Pr(<", letter, ")"),
    greater=paste0("Pr(>", letter, ")"),
    two.sided=paste0("Pr(>|", letter, "|)"))
colnames(mtests) = c("Estimate", "Std. Error", paste(letter, "value"), pname)
alt = switch(comparisons$alternative, two.sided="==", less=">=", greater="<=")
rownames(mtests) = paste(rownames(mtests), alt, comparisons$rhs)
kable(mtests, digits=3)
```
